Equations & R Code

### 2016-11-27 manuscript/coexistence-stability_ms.Rmd
finalizes PBA edits

\Large{How fluctuation-dependent species coexistence affects the diversity-stability relationship}

ALTERNATIVE TITLES:
-FLUCTUATION-DEPENDENT COEXISTENCE AND THE DIVERSITY-STABILITY RELATIONSHIP
-THE DIVERSITY-STABILITY RELATIONSHIP IS SHAPED BY THE MECHANISM OF COEXISTENCE

@@

@@

However, the ability of biodiversity-ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
Much of the debate centers around the fact BEF experimental protocols do not allow species gains to offset species losses.
Theoretical work on diversity-stability relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].
THE CHANGES I MADE IN THE PREVIOUS PARAGRAPH ARE REDUNDANT WITH THE END OF THE NEXT PARAGRAPH. YOU MIGHT NEED TO REORGANIZE, BUT I THOUGHT IT WAS IMPORTANT TO GET TO THESE POINTS SOONER THAN LATER.

@@

To analyze the above results in more detail [THIS IS VAGUE, WHAT WAS THE OBJECTIVE?], we simulated our model over a range of environmental variance and species pool sizes.
In Fig. 3 we show storage effect simulation results where environmental variability determines species coexistence from a regional species pool of four species.
We also show results from nested subsets of the four species pool (e.g., only two species in the pool instead of four) to show the trajectory of ecosystem variability if new species are not present to join the local community.
I FIND THE PREVIOUS TWO SENTENCES CONFUSING and METHODS-Y. CUT?
Realized species richness increased with environmental variability and, in some cases, increasing variability completely offset the effect of increasing environmental variability on stability.
More species rich communities were less variable on average and increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig. S1-3).

@@

When coexistence is fluctuation-dependent, however, the outcome is less certain.
By simulating communities with different species pool sizes across a gradient of environmental variabilitym, we showed that species gains due to increasing environmental variability can buffer the direct effect of environmental variability on ecosystem variability (Figs. 3 and 4).
These results lead to two conclusions.

@@

Species coexistence in real ecological communities probably emerges from some combination of fluctuation-independent and fluctuation-dependent mechanisms [@Chesson2000; @Clark2010].
Likewise, environmental conditions in real ecosystems are unlikely to change only in their variability without an associated change in the mean [@Avolio2015].
Therefore, environmental change has the potential to alter the niche and fitness differences among species in multiple ways, some of which were not present in our current analysis.
Mean changes in environmental conditions could reorder competitive hierarchies [@Klanderud2005] and/or alter the availability of niches [@Harpole2016].
Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species' response to the particular environmental driver, and biotic interactions [@Hallett2014a].
Thus, it is becoming clear that understanding how ecosystem stability will respond to global change will require a trait-based approach.

I DON'T LIKE THIS CONCLUSION MUCH. THE ARGUMENT FOR A TRAIT BASED APPROACH SEEMS PRETTY SUPERFICIAL AND CHANGES THE SUBJECT FROM OUR MAIN FOCUS. BUT I LIKE THE CAVEAT REGARDING SHIFTS IN MEANS ALONG WITH VARIANCES. COULD THAT BE MOVED UP, AND WE END WITH THE PREVIOUS PARAGRAPH, WHICH IS STRONG?

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=5in]{./components/storage_effect_div+envar_varycomp.png}
  \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via the storage effect. Panels (A-C) show simulation results where species have slightly asymmetrical competitive effects, whereas panels (D-F) show results when competition is more asymmetric. We show results for different levels of correlations of species' environmental responses, $\rho$.}
\end{figure}

>>>>

\Large{The diversity-stability relationship is shaped by the mechanism of coexistence}

@@

<!--
ALTERNATIVE TITLES:
-FLUCTUATION-DEPENDENT COEXISTENCE AND THE DIVERSITY-STABILITY RELATIONSHIP
-THE DIVERSITY-STABILITY RELATIONSHIP IS SHAPED BY THE MECHANISM OF COEXISTENCE
-How fluctuation-dependent species coexistence affects the diversity-stability relationship
-->

@@

However, the ability of biodiversity-ecosystem functioning (BEF) experiments to accurately represent real-world dynamics is debated [@Wardle2016; @Eisenhauer2016].
Much of the debate centers around the fact BEF experimental protocols do not allow species gains to offset species losses.
Theoretical work on diversity-stability relationships typically suffers from the same limitation: it recognizes the role of environmental variability in driving population fluctuations which destabilize ecosystems, but ignores the potential for environmental variability to promote species richness and thereby help stabilize ecosystems [Loreau 2010, but see @Chesson2001].

@@

To understand how much species additions can stabilize ecosystem functioning, we simulated our model over a range of environmental variance and species pool sizes.
Realized species richness increased with environmental variability and, in some cases, increasing variability completely offset the effect of increasing environmental variability on stability.
More species rich communities were less variable on average and increased in ecosystem *CV* at a slower rate than communities with fewer species (e.g., lower slopes in log-log space; Fig. S1-3).

@@

When coexistence is fluctuation-dependent, however, the outcome is less certain.
By simulating communities with different species pool sizes across a gradient of environmental variability, we showed that species gains due to increasing environmental variability can buffer the direct effect of environmental variability on ecosystem variability (Figs. 3 and 4).
These results lead to two conclusions.

@@

<!--
Species coexistence in real ecological communities probably emerges from some combination of fluctuation-independent and fluctuation-dependent mechanisms [@Chesson2000; @Clark2010].
Likewise, environmental conditions in real ecosystems are unlikely to change only in their variability without an associated change in the mean [@Avolio2015].
Therefore, environmental change has the potential to alter the niche and fitness differences among species in multiple ways, some of which were not present in our current analysis.
Mean changes in environmental conditions could reorder competitive hierarchies [@Klanderud2005] and/or alter the availability of niches [@Harpole2016].
Associated changes in ecosystem stability will depend upon the magnitude of environmental change, each species' response to the particular environmental driver, and biotic interactions [@Hallett2014a].
-->

@@

\begin{figure}[!ht]
  \centering
      \includegraphics[height=5in]{./components/storage_effect_div+envar_varycomp.png}
  \caption{The effect of environmental variability on ecosystem variability with associated effects of species richness when species coexist via the storage effect. Panels (A-C) show simulation results where species have slightly asymmetrical competitive effects, whereas panels (D-F) show results when competition is more asymmetric. We show results for different levels of correlations of species' environmental responses, $\rho$. Colored vertical lines show the level of environmental variability at which each level of species richness first occurs.}
\end{figure}

### 2016-11-27 manuscript/coexistence-stability_ms.tex
finalizes PBA edits

However, the ability of biodiversity-ecosystem functioning (BEF)
experiments to accurately represent real-world dynamics is debated
(Wardle 2016, Eisenhauer et al. 2016). Much of the debate centers around
the fact BEF experimental protocols do not allow species gains to offset
species losses. Theoretical work on diversity-stability relationships
typically suffers from the same limitation: it recognizes the role of
environmental variability in driving population fluctuations which
destabilize ecosystems, but ignores the potential for environmental
variability to promote species richness and thereby help stabilize
ecosystems (Loreau 2010, but see Chesson et al. 2001). THE CHANGES I
MADE IN THE PREVIOUS PARAGRAPH ARE REDUNDANT WITH THE END OF THE NEXT
PARAGRAPH. YOU MIGHT NEED TO REORGANIZE, BUT I THOUGHT IT WAS IMPORTANT
TO GET TO THESE POINTS SOONER THAN LATER.

@@

To analyze the above results in more detail {[}THIS IS VAGUE, WHAT WAS
THE OBJECTIVE?{]}, we simulated our model over a range of environmental
variance and species pool sizes. In Fig. 3 we show storage effect
simulation results where environmental variability determines species
coexistence from a regional species pool of four species. We also show
results from nested subsets of the four species pool (e.g., only two
species in the pool instead of four) to show the trajectory of ecosystem
variability if new species are not present to join the local community.
I FIND THE PREVIOUS TWO SENTENCES CONFUSING and METHODS-Y. CUT? Realized
species richness increased with environmental variability and, in some
cases, increasing variability completely offset the effect of increasing
environmental variability on stability. More species rich communities
were less variable on average and increased in ecosystem \emph{CV} at a
slower rate than communities with fewer species (e.g., lower slopes in
log-log space; Fig. S1-3).

@@

Species coexistence in real ecological communities probably emerges from
some combination of fluctuation-independent and fluctuation-dependent
mechanisms (Chesson 2000a, Clark et al. 2010). Likewise, environmental
conditions in real ecosystems are unlikely to change only in their
variability without an associated change in the mean (Avolio et al.
2015). Therefore, environmental change has the potential to alter the
niche and fitness differences among species in multiple ways, some of
which were not present in our current analysis. Mean changes in
environmental conditions could reorder competitive hierarchies
(Klanderud and Totland 2005) and/or alter the availability of niches
(Harpole et al. 2016). Associated changes in ecosystem stability will
depend upon the magnitude of environmental change, each species'
response to the particular environmental driver, and biotic interactions
(Hallett et al. 2014). Thus, it is becoming clear that understanding how
ecosystem stability will respond to global change will require a
trait-based approach.

I DON'T LIKE THIS CONCLUSION MUCH. THE ARGUMENT FOR A TRAIT BASED
APPROACH SEEMS PRETTY SUPERFICIAL AND CHANGES THE SUBJECT FROM OUR MAIN
FOCUS. BUT I LIKE THE CAVEAT REGARDING SHIFTS IN MEANS ALONG WITH
VARIANCES. COULD THAT BE MOVED UP, AND WE END WITH THE PREVIOUS
PARAGRAPH, WHICH IS STRONG?

@@

Avolio, M. L., K. J. L. Pierre, G. R. Houseman, S. E. Koerner, E. Grman,
F. Isbell, D. S. Johnson, and K. R. Wilcox. 2015. A framework for
quantifying the magnitude and variability of community responses to
global change drivers. Ecosphere 6:1--14.

@@

Clark, J. S., D. Bell, C. Chu, B. Courbaud, M. Dietze, M. Hersh, J.
HilleRisLambers, I. Ib{á}{ñ}ez, S. LaDeau, S. McMahon, J. Metcalf, J.
Mohan, E. Moran, L. Pangle, S. Pearson, C. Salk, Z. Shen, D. Valle, and
P. Wyckoff. 2010. High-dimensional coexistence based on individual
variation: a synthesis of evidence. Ecological Monographs 80:569--608.

@@

Hallett, L. M., J. S. Hsu, E. E. Cleland, S. L. Collins, T. L. Dickson,
E. C. Farrer, L. A. Gherardi, K. L. Gross, R. J. Hobbs, L. Turnbull, and
K. N. Suding. 2014. Biotic mechanisms of community stability shift along
a precipitation gradient. Ecology 95:1693--1700.

Harpole, W. S., L. L. Sullivan, E. M. Lind, J. Firn, P. B. Adler, E. T.
Borer, J. Chase, P. A. Fay, Y. Hautier, H. Hillebrand, A. S. MacDougall,
E. W. Seabloom, R. Williams, J. D. Bakker, M. W. Cadotte, E. J.
Chaneton, C. Chu, E. E. Cleland, C. D'Antonio, K. F. Davies, D. S.
Gruner, N. Hagenah, K. Kirkman, J. M. H. Knops, K. J. {La Pierre}, R. L.
McCulley, J. L. Moore, J. W. Morgan, S. M. Prober, A. C. Risch, M.
Schuetz, C. J. Stevens, and P. D. Wragg. 2016. Addition of multiple
limiting resources reduces grassland diversity. Nature 537:93--96.

@@

Klanderud, K., and Ø. Totland. 2005. Simulated climate change altered
dominance hierarchies and diversity of an alpine biodiversity hotspot.
Ecology 86:2047--2054.

>>>>

However, the ability of biodiversity-ecosystem functioning (BEF)
experiments to accurately represent real-world dynamics is debated
(Wardle 2016, Eisenhauer et al. 2016). Much of the debate centers around
the fact BEF experimental protocols do not allow species gains to offset
species losses. Theoretical work on diversity-stability relationships
typically suffers from the same limitation: it recognizes the role of
environmental variability in driving population fluctuations which
destabilize ecosystems, but ignores the potential for environmental
variability to promote species richness and thereby help stabilize
ecosystems (Loreau 2010, but see Chesson et al. 2001).

@@

To understand how much species additions can stabilize ecosystem
functioning, we simulated our model over a range of environmental
variance and species pool sizes. Realized species richness increased
with environmental variability and, in some cases, increasing
variability completely offset the effect of increasing environmental
variability on stability. More species rich communities were less
variable on average and increased in ecosystem \emph{CV} at a slower
rate than communities with fewer species (e.g., lower slopes in log-log
space; Fig. S1-3).

@@

@@

@@

@@

@@
